---
title: "MSDS 6306 - Doing Data Science - Tidy Data"
author: "Randall Hendrickson"
date: "October 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Live Session Unit 05 assignment - Tidy Data

```{r global_data}
library(RCurl)


github_file_1 <- "https://raw.githubusercontent.com/rhendrickson42/msds6306/master/homework/livesession05/data/yob2016.txt"
github_local_1 <- "./yob2016-github.txt"

github_file_2 <- "https://raw.githubusercontent.com/rhendrickson42/msds6306/master/homework/livesession05/data/yob2015.txt"
github_local_2 <- "./yob2015-github.txt"

# download files from Github for munging
download.file(github_file_1, destfile = github_local_1, method = "curl")
download.file(github_file_2, destfile = github_local_2, method = "curl")

```



### 1. Data Munging



```{r data_munging}


df <- read.table(github_local_1, header=FALSE, sep=";", na.strings="NA", dec=".", strip.white=TRUE)

# read from local filesystem if github is not working
# df <- read.table("D:/dataScience/doingDS/hw05/yob2016.txt", header=FALSE, sep=";", na.strings="NA", dec=".", strip.white=TRUE)

names(df)[c(1,2,3)] <- c("First_Name","Gender","Amount_Of_Children_2016")

# display the summary and structure of df

#summary of df
summary(df)

# str of df
str(df)

# c. find name yyy
# $ anchors end of line
df[grep("yyy$", df$First_Name),]

# remove observation from above
y2016 <- df[-grep("yyy$", df$First_Name),]

```

### 2. Data Merging


```{r data_merging, echo=TRUE}


# Question 2

y2015 <- read.table(github_local_2, header=FALSE, sep=",")
names(y2015)[c(1,2,3)] <- c("First_Name","Gender","Amount_Of_Children_2015")

# display the last ten rows
# describe something interesting about these 10 rows

tail(y2015, n = 10L)

```

#### Describe something interesting about these last 10 rows

These 10 rows are all male names that were used 5 times each in 2015

Merge y2016 and y2015 by your Name column; assign it to final. 
The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.


```{r data_merging_2, echo=TRUE}

final <- merge(y2015, y2016, by=c("First_Name", "Gender"))
# make sure no NAs
sapply(final, function(x) sum(is.na(x)))

# show last ten rows in the dataframe
#tail(final, n = 10L)

```


### 3. Data Summary

Create a new column called Total that adds the amount of children in 2015 and 2016 together.

```{r data_summary}

# final$Total <- "Total"
final$Total <- final$Amount_Of_Children_2015 + final$Amount_Of_Children_2016

# a. how many people were given popular names
total_popular_names <- sum(final$Total)
total_popular_names


# b. 

top10 <- head(final[order(final$Total, decreasing = T),], n = 10)
top10 


# First_Name Gender Amount_Of_Children_2015 Amount_Of_Children_2016 Total
# 8290        Emma      F                   20415                   19414 39829
# 19886     Olivia      F                   19638                   19246 38884
# 19594       Noah      M                   19594                   19015 38609
# 16114       Liam      M                   18330                   18138 36468
# 23273     Sophia      F                   17381                   16070 33451
# 3252         Ava      F                   16340                   16237 32577
# 17715      Mason      M                   16591                   15192 31783
# 25241    William      M                   15863                   15668 31531
# 10993      Jacob      M                   15914                   14416 30330
# 10682   Isabella      F                   15574                   14722 30296


# c. the client is expecting a girl

#girl_names <- filter(final, Gender == "F")
girl_names <- subset(final, Gender =='F')

girl <- head(girl_names[order(girl_names$Total, decreasing = T),], n = 10)

girl_subset <- girl[c("First_Name", "Total")]
write.csv(girl_subset, file = "itsagirl.csv", row.names = FALSE)

girl_subset

```

### 4. Upload to GitHub

Here is the location of the GitHub project

https://github.com/rhendrickson42/msds6306/tree/master/homework/livesession05

